\usepackage{etex}
\usepackage[letterpaper, inner=1in,outer=1in,top=1in,bottom=1in]{geometry}
\usepackage{csquotes}
\usepackage{mathtools,amsfonts,amssymb,amsthm}
\usepackage{empheq}

\usepackage[american]{babel}
% Uncomment the next two lines if a bibliography needs to be included.
\usepackage[backend=biber, style=apa]{biblatex}
\addbibresource{References.bib}

\usepackage[none]{hyphenat}
\raggedright

\usepackage[utf8]{inputenc} % For unicode characters
\usepackage[T1]{fontenc} % For unicode characters
\usepackage{fontspec} % For imported font files
\usepackage{titling
}
\usepackage[doublespacing]{setspace}
\usepackage{blindtext}
\usepackage[dvipsnames]{xcolor}
\usepackage{listings}
\usepackage{gensymb}
\usepackage{enumitem}
\usepackage{wrapfig}
\usepackage{graphicx}
\usepackage[linesnumbered,ruled,lined,commentsnumbered]{algorithm2e}
\usepackage{pdfpages}
\usepackage{qtree}
\usepackage{xcolor}
\usepackage{float}
\usepackage{textpos}[absolute]
\usepackage{lipsum}
%\usepackage{tipa}
% Keep hyperref as the last package.
\usepackage{hyperref}

{\IfFontExistsTF{CourierNew}{\newfontfamily\CodeBlock{CourierNew}}{\newfontfamily\CodeBlock{CharisSILR.ttf}}
{\IfFontExistsTF{Calibri}{\newfontfamily\PhoneticFont{Calibri}}{\newfontfamily\PhoneticFont{CharisSILR.ttf}}


%\renewcommand{\labelenumii}{\arabic{enumi}.\arabic{enumii}}
%\renewcommand{\labelenumiii}{\arabic{enumi}.\arabic{enumii}.\arabic{enumiii}}
%\renewcommand{\labelenumiv}{\arabic{enumi}.\arabic{enumii}.\arabic{enumiii}.\arabic{enumiv}}

% Set up the Functional Requirements List
\newlist{functionalRequirements}{enumerate}{5}
\setlist[functionalRequirements, 1]{label=FR.\arabic{functionalRequirementsi}., ref=FR.\arabic{functionalRequirementsi}}
\setlist[functionalRequirements, 2]{label=FR.\arabic{functionalRequirementsi}.\arabic{functionalRequirementsii}.,
  ref=FR.\arabic{functionalRequirementsi}.\arabic{functionalRequirementsii}}
\setlist[functionalRequirements, 3]{label=FR.\arabic{functionalRequirementsi}.\arabic{functionalRequirementsii}.\arabic{functionalRequirementsiii}.,
  ref=FR.\arabic{functionalRequirementsi}.\arabic{functionalRequirementsii}.\arabic{functionalRequirementsiii}}
\setlist[functionalRequirements, 4]{label=FR.\arabic{functionalRequirementsi}.\arabic{functionalRequirementsii}.\arabic{functionalRequirementsiii}.\arabic{functionalRequirementsiv}.,
  ref=FR.\arabic{functionalRequirementsi}.\arabic{functionalRequirementsii}.\arabic{functionalRequirementsiii}.\arabic{functionalRequirementsiv}}
\setlist[functionalRequirements, 5]{label=FR.\arabic{functionalRequirementsi}.\arabic{functionalRequirementsii}.\arabic{functionalRequirementsiii}.\arabic{functionalRequirementsiv}.\arabic{functionalRequirementsv}.,
  ref=FR.\arabic{functionalRequirementsi}.\arabic{functionalRequirementsii}.\arabic{functionalRequirementsiii}.\arabic{functionalRequirementsiv}.\arabic{functionalRequirementsv}}

% Set up the Non-functional Requirements List
\newlist{nonFunctionalRequirements}{enumerate}{5}
\setlist[nonFunctionalRequirements, 1]{label=NR.\arabic{nonFunctionalRequirementsi}.,
  ref=NR.\arabic{nonFunctionalRequirementsi}}
\setlist[nonFunctionalRequirements, 2]{label=NR.\arabic{nonFunctionalRequirementsi}.\arabic{nonFunctionalRequirementsii}.,
  ref=NR.\arabic{nonFunctionalRequirementsi}.\arabic{nonFunctionalRequirementsii}}
\setlist[nonFunctionalRequirements, 3]{label=NR.\arabic{nonFunctionalRequirementsi}.\arabic{nonFunctionalRequirementsii}.\arabic{nonFunctionalRequirementsiii}.,
  ref=NR.\arabic{nonFunctionalRequirementsi}.\arabic{nonFunctionalRequirementsii}.\arabic{nonFunctionalRequirementsiii}}
\setlist[nonFunctionalRequirements, 4]{label=NR.\arabic{nonFunctionalRequirementsi}.\arabic{nonFunctionalRequirementsii}.\arabic{nonFunctionalRequirementsiii}.\arabic{nonFunctionalRequirementsiv}.,
  ref=NR.\arabic{nonFunctionalRequirementsi}.\arabic{nonFunctionalRequirementsii}.\arabic{nonFunctionalRequirementsiii}.\arabic{nonFunctionalRequirementsiv}}
\setlist[nonFunctionalRequirements, 5]{label=NR.\arabic{nonFunctionalRequirementsi}.\arabic{nonFunctionalRequirementsii}.\arabic{nonFunctionalRequirementsiii}.\arabic{nonFunctionalRequirementsiv}.\arabic{nonFunctionalRequirementsv}.,
  ref=NR.\arabic{nonFunctionalRequirementsi}.\arabic{nonFunctionalRequirementsii}.\arabic{nonFunctionalRequirementsiii}.\arabic{nonFunctionalRequirementsiv}.\arabic{nonFunctionalRequirementsv}}
  
 % algorithm2e definitions and settings
\SetKwProg{Alg}{}{}{}
\SetKwProg{Fn}{Function}{ is}{end}
\SetKwProg{Tst}{Test}{::}{end}
\SetKwBlock{KWLambda}{Lambda}{end}
\SetKwInOut{Input}{input}
\SetKwInOut{Output}{output}
\SetKwInOut{Exception}{exception}
\SetKw{Enum}{enum}
\SetKw{Const}{constant}
\SetKw{Event}{event}
\SetKwProg{SplitFnStart}{Function}{ is}{}
\SetKwBlock{SplitFnMid}{}{}
\SetKwBlock{SplitFnEnd}{}{end}
\SetKwRepeat{Do}{do}{while}
\SetKwProg{Try}{try}{:}{}{}
\SetKwProg{Catch}{catch}{:}{end}
\SetKwProg{uCatch}{catch}{:}{}
\SetKwProg{Finally}{finally}{:}{end}
% End algorithm2e definitions and settings

%% Workaround for APA issue with director citation (but not APA6)
\renewbibmacro*{apaeditorstrg}[1]{%
  \iffieldundef{#1type}
    {\ifthenelse{\value{#1}>1\OR\ifandothers{#1}}
       {\bibcpstring{editors}}
       {\bibcpstring{editor}}}
    {\ifthenelse{\value{#1}>1\OR\ifandothers{#1}}
       {\ifbibstring{type\thefield{#1type}s}
          {\bibcpstring{type\thefield{#1type}s}}
          {\bibcpstring{\thefield{#1type}s}}}
       {\ifbibstring{type\thefield{#1type}}
          {\bibcpstring{type\thefield{#1type}}}
          {\bibcpstring{\thefield{#1type}}}}}}
%% End APA citation workaround

% hyperref setup
\hypersetup{
    colorlinks=true,
    linkcolor=blue,
    filecolor=magenta,      
    urlcolor=cyan,
    citecolor=blue
    }

\urlstyle{same}
% End hyperref setup

% Replace with your name
\author{Ronald B. Oakes\\ ro421@mynsu.nova.edu}
